---
title: "Variants detected by Strelka2 in cNF cell lines"
author: Robert Allaway
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: paged 
    #code_folding: true 
---

# Setup

First, we set up our workspace. Load maftools to analyze the maf files.

Get all the mafs and merge them into one big maf.

```{r message=FALSE, warning=FALSE}
library(maftools)
library(tidyverse)
library(synapser)
synLogin()

child <- synGetChildren('syn35817218')$asList()
ids <- map(child, "id") %>% unlist()

maf_paths <- lapply(ids, synGet)

mafs <- lapply(maf_paths, function(x){
  maftools::read.maf(x$path)
})

merged_maf <- merge_mafs(mafs)


```

# Compare mafs

First, let's compare the immortalized vs primary cells to see if there are any significant differentially mutated genes. 
There do not appear to be (lowest p-value ~0.1 and all p-values = 1 after FDR correction). With more cell lines, we might be able to detect a difference here, but with the current experiment we don't detect any significantly differentially mutated genes. 

```{r message=FALSE, warning=FALSE}

primary_mafs <- subsetMaf(merged_maf, tsb = c("28cNF","cNF00.10a","cNF04.9a","cNF97.2a","cNF97.2b","cNF98.4c","cNF98.4d"))

immortalized_mafs <- subsetMaf(merged_maf, tsb = c("i28cNF","icNF00.10a","icNF04.9a","icNF97.2a","icNF97.2b","icNF98.4c","icNF98.4d"))

compare_mafs <- mafCompare(primary_mafs,immortalized_mafs, m1Name = "Primary", m2Name = "Immortalized", minMut = 0)

compare_mafs[1]
```


# Plot maf summaries

Let's plot a summary of each maf. 

```{r message=FALSE, warning=FALSE}

lapply(mafs, function(x){
       plotmafSummary(x)
  })

```

# Plot maf summaries

Let's plot a summary of each _cohort_.

```{r message=FALSE, warning=FALSE}

p <- plotmafSummary(primary_mafs, showBarcodes = T, textSize = 0.5)
p

pdf('../figures/Strelka2_primaryMafSummary.pdf')
p
dev.off()

p <- plotmafSummary(immortalized_mafs, showBarcodes = T, textSize = 0.5)
p

pdf('../figures/Strelka2_immortalizedMafSummary.pdf')
p
dev.off()

```

## Look at NF1 variants

```{r message=FALSE, warning=FALSE}

p <- lollipopPlot2(primary_mafs, immortalized_mafs, gene = 'NF1',m1_label = 'all', m2_label = 'all',labPosAngle = 45, labPosSize = 0.75, domainLabelSize = 0.5, m1_name = "Primary", m2_name = "Immortalized")

p

pdf("../figures/DeepVariant_NF1_variants.pdf")
p
dev.off()


```


## Table of NF1 variants

```{r message=FALSE, warning=FALSE}

merged_maf@data %>% 
  filter(Hugo_Symbol == "NF1") %>% 
  select(Tumor_Sample_Barcode, Variant_Classification, Variant_Type, 
         HGVSc, HGVSp_Short,
         dbSNP_RS, SIFT, PolyPhen, IMPACT) %>% 
  set_names(c("Cell Line", "Variant Classification", "Variant Type",
          "Genetic Change", "Protein Change", "dbSNP ID", "SIFT", "PolyPhen", "IMPACT")) %>% 
  write_csv("../figures/Strelka2_NF1_variants.csv")

```

```{r message=FALSE, warning=FALSE}

sessionInfo()

```